**OptiFit: a fast method for fitting amplicon sequences to existing OTUs**

### 2021-11-19 paper/manuscript.Rmd
undo change equally to similarly

These results demonstrate a Goldilocks effect such that consideration of the appropriate taxonomic resolution for utilizing the microbiome as a predictive tool is warranted. In general, we found that finer taxonomic resolution (e.g. ASV) did not add additional sensitivity to predicting SRNs based on microbiome composition. Family, genus, and OTU level data all performed similarly. At the ASV level the fine resolution actually impeded model performance due to the sparsity of shared taxa and led to decreased model performance. The tendency for ASV level annotation to split single bacterial genomes into multiple taxa [@schloss] could also be a contributing factor to the sparsity of shared taxa. Additionally, these results indicate that there are not specific individual bacterial strains that are useful to resolve SRNs, rather sets of closely related bacterial taxa. Overall, either family, genus, or OTU level taxonomy appear to perform similarly for predicting subjects with SRNs based on the composition of the gut microbiome. A potential benefit of utilizing genus or family level data could be that it may allow for merging data generated from different 16S rRNA gene regions or sequencing platforms.

>>>>

These results demonstrate a Goldilocks effect such that consideration of the appropriate taxonomic resolution for utilizing the microbiome as a predictive tool is warranted. In general, we found that finer taxonomic resolution (e.g. ASV) did not add additional sensitivity to predicting SRNs based on microbiome composition. Family, genus, and OTU level data all performed equally At the ASV level the fine resolution actually impeded model performance due to the sparsity of shared taxa and led to decreased model performance. The tendency for ASV level annotation to split single bacterial genomes into multiple taxa [@schloss] could also be a contributing factor to the sparsity of shared taxa. Additionally, these results indicate that there are not specific individual bacterial strains that are useful to resolve SRNs, rather sets of closely related bacterial taxa. Overall, either family, genus, or OTU level taxonomy appear to perform equally for predicting subjects with SRNs based on the composition of the gut microbiome. A potential benefit of utilizing genus or family level data could be that it may allow for merging data generated from different 16S rRNA gene regions or sequencing platforms.

### 2021-11-19 paper/manuscript.tex
undo change equally to similarly

\({^\#}\) Current Affiliation: Exploratory Science Center, Merck \& Co.,
Inc., Cambridge, Massachusetts, USA~

@@

These results demonstrate a Goldilocks effect such that consideration of
the appropriate taxonomic resolution for utilizing the microbiome as a
predictive tool is warranted. In general, we found that finer taxonomic
resolution (e.g.~ASV) did not add additional sensitivity to predicting
SRNs based on microbiome composition. Family, genus, and OTU level data
all performed equally. At the ASV level the fine resolution actually
impeded model performance due to the sparsity of shared taxa and led to
decreased model performance. The tendency for ASV level annotation to
split single bacterial genomes into multiple taxa (12) could also be a
contributing factor to the sparsity of shared taxa. Additionally, these
results indicate that there are not specific individual bacterial
strains that are useful to resolve SRNs, rather sets of closely related
bacterial taxa. Overall, either family, genus, or OTU level taxonomy
appear to perform equally for predicting subjects with SRNs based on the
composition of the gut microbiome. A potential benefit of utilizing
genus or family level data could be that it may allow for merging data
generated from different 16S rRNA gene regions or sequencing platforms.

@@

\includegraphics[width=\textwidth,height=0.45\textheight]{figure_s3.png}\\
\textbf{Supplemental Figure 3: Prevalence of Taxa in Samples.}
Distribution of the prevalence of taxa across samples at each taxonomic
level. Percent of samples is split into 10 groups where the first is for
taxa present in 0 to 10\% of samples, then \textgreater10\% to 20\% of
samples, and so on. The total number of taxa for each taxonomic level
after preprocessing is in parenthesis next to the title of the plot (or
the name of the taxonomic level).

@@

\includegraphics[width=\textwidth,height=0.85\textheight]{figure_s4.png}
\textbf{Supplemental Figure 4: Top 10 important taxa at each taxonomic
level.} Summary of the 10 most important taxa for the random forest
models at each taxonomic level based on the average decrease in AUROC
when the feature is permuted. Dot represents the mean decrease in AUROC
and the lines extending from the dot represent the standard deviation
from the mean.

>>>>

\({^\#}\) Current Affiliation: Bristol Myers Squibb, Summit, New Jersey,
USA~

@@

These results demonstrate a Goldilocks effect such that consideration of
the appropriate taxonomic resolution for utilizing the microbiome as a
predictive tool is warranted. In general, we found that finer taxonomic
resolution (e.g.~ASV) did not add additional sensitivity to predicting
SRNs based on microbiome composition. Family, genus, and OTU level data
all performed equally At the ASV level the fine resolution actually
impeded model performance due to the sparsity of shared taxa and led to
decreased model performance. The tendency for ASV level annotation to
split single bacterial genomes into multiple taxa (12) could also be a
contributing factor to the sparsity of shared taxa. Additionally, these
results indicate that there are not specific individual bacterial
strains that are useful to resolve SRNs, rather sets of closely related
bacterial taxa. Overall, either family, genus, or OTU level taxonomy
appear to perform equally for predicting subjects with SRNs based on the
composition of the gut microbiome. A potential benefit of utilizing
genus or family level data could be that it may allow for merging data
generated from different 16S rRNA gene regions or sequencing platforms.

@@

\includegraphics[width=\textwidth,height=0.45\textheight]{figure_s3.png}\\
\textbf{Supplemental Figure 3: Prevalence of Taxa in Samples.}
Distribution of the prevalence of taxa across samples at each taxonomic
level. Percent of samples is split into 10 groups where the first is for
taxa present in 0 to 10\% of samples, then \textgreater10\% to 20\% of
samples, and so on. The total number of taxa for each taxonomic level
after preprocessing is in parenthesis next to the title of the plot.

@@

\includegraphics[width=\textwidth,height=0.85\textheight]{figure_s4.png}
\textbf{Supplemental Figure 4: Top 10 important taxa at each taxonomic
level.} Summary of the 10 most important taxa for the random forest
models at each taxonomic level based on the average decrease in AUROC
when the feature is permuted. Dot represents the mean decrease in AUROC
and the lines extending from the dot represent the standard deviation
from the mean. Taxonomic annotations are only reliably assigned to the
genus level, therefore the OTU and ASV plots have genus labels.
